Mining HIV Trends in Social Media Data

### 2016-04-13 draft/draft.tex
too tired to continue today

High level, but more specific than intro, introduce the approach and methods used in this paper here.

@@

\begin{table}
\centering
\caption{Cosine Similarity to 'prep'}
\begin{tabular}{|l|c|} \hline
Related Word & Cosine Similarity to 'prep'\\ \hline
truvada & 0.836507 \hline
benegative & 0.809185\\ \hline
charliesheen & 0.805214\\ \hline
hivtestweek & 0.797361\\ \hline
sexwork & 0.774867\\ \hline
gettested & 0.765772\\ \hline
hiv & 0.765389\\ \hline
icasa2015 & 0.749948\\ \hline
doingit & 0.749230\\ \hline
aids & 0.738572\\ \hline
\hline\end{tabular}
\end{table}

\begin{table}
\centering
\caption{Cosine Similarity to 'truvada'}
\begin{tabular}{|c|c|} \hline
Related Word & Cosine Similarity to 'truvada'\\ \hline
truvada & 0.836507 \hline
benegative & 0.809185\\ \hline
charliesheen & 0.805214\\ \hline
hivtestweek & 0.797361\\ \hline
sexwork & 0.774867\\ \hline
gettested & 0.765772\\ \hline
hiv & 0.765389\\ \hline
icasa2015 & 0.749948\\ \hline
doingit & 0.749230\\ \hline
aids & 0.738572\\ \hline
\hline\end{tabular}
\end{table}

@@

\begin{figure}
\centering
\includegraphics[height=1in, width=1in]{fly}
\caption{Time domain plot.}
\end{figure}

@@

\subsubsection{Sentiment Classification}

@@

Acknowledgements go here.

>>>>

We sought to determine trends in HIV and PrEP discourse on twitter, to inform and coordinate public health efforts to promote PrEP adoption and adherence for at-risk individuals. We collected 624,569 tweets related to 'HIV', 'AIDS', 'truvada', 'prophylaxis', 'imtesting', 'PrEP' from Twitter's streaming API. The tweets were restricted to English language, and the collection dates spanned from the 47th week of 2015 to the 14th week of 2016. The tweets were and cleaned of exotic characters.

% Prior to topic modelling, tweets were cleaned of high and low frequency words, and transformed by Term Frequency Inverse Document Frequency (TFIDF) processing.

We excluded non-English tweets, and also tweets that did not originate in Contiguous United States time zones. A subset of the tweets we collected had geolocation data available. We found that tweets were largely concentrated on US and Canadian metro areas (Figure 1).

\begin{figure}
\centering
\includegraphics[height=2.5in, width=3.5in]{map}
\caption{Plot of geolocated tweets.}
\end{figure}

@@

% show text of relevent tweets from d2v

% talk a bit about pre processing

The first analyses that we performed sought to identify certain keywords, hashtags, tweets and users, that were discussing HIV and PrEP related contexts. Word2Vec and the related method, Paragraph2Vec are unsupervised machine learning methods that have performed well at embedding natural language in a semantic vector space. In our analysis, Word2Vec allows us to determine semantically similar words, while Paragraph2Vec allows us to determine similar tweets, users and hashtags.

% table 1 - w2v PrEP
\begin{table}
\centering
\caption{Cosine Similarity to 'truvada'}
\begin{tabular}{|l|c|} \hline
Related Word & Cosine Similarity to 'truvada'\\ \hline
truvada & 0.796666\\ \hline
Truvada & 0.793999\\ \hline
DoingIt & 0.738141\\ \hline
WorldAidsDay & 0.728808\\ \hline
WorldAIDSDay & 0.720910\\ \hline
NancyReagan & 0.717667\\ \hline
NBHAAD & 0.705061\\ \hline
ART & 0.704300\\ \hline
hiv & 0.702844\\ \hline
HLM2016AIDS & 0.698698\\ \hline
\hline\end{tabular}
\end{table}

% table 2 - d2v PrEP
\begin{table}
\centering
\caption{Cosine Similarity to '\#truvada'}
\begin{tabular}{|l|c|} \hline
Related Hashtag/Tweet & Cosine Similarity to '\#truvada'\\ \hline
\#lgbtmedia16 & 0.739128\\ \hline
\#hiv & 0.727602\\ \hline
\#whereisprep & 0.707165\\ \hline
\#truvada & 0.696113\\ \hline
\#hivprevention & 0.636068\\ \hline
702179860983189504 & 0.630055\\ \hline
user-711275699529764864 & 0.629254\\ \hline
708519265540907010 & 0.628778\\ \hline
712032637024653313 & 0.628646\\ \hline
\#harrogatehour & 	0.628547\\ \hline
\hline\end{tabular}
\end{table}

% show PCA for each table above (remember to mention percent varience)

@@

% PrEP related trends (2 topic time series worth)
% World AIDs day as validation (1 topic time series)

Next we sought to identify some temporal trends in PrEP related trends. We used Dynamic Topic Modelling to identify how certain topics change over time. We specified 10 topics and plotted representative words from two of the topics.

In topic 3 we see that the keyword 'prep' remains relatively constant, while 'prevention' and 'truvada' decline over time. 'hivaids' increases and then decreases in magnitude over the course of the time series (Figure 2). These dynamics may indicate short term events in hivaids and PrEP-related conversation on twitter over the time period studied.

We found that topic 5 captured several keywords related to World AIDS Day (Figure 3). We can see that all of these terms peak in the 47th week of 2015 and then decline into 2016. This correlates well with the actual date of World AIDS Day, December 1st. While we aren't specifically interested in World AIDS Day to inform our understanding of PrEP discussion, this observation validates our ability to identify temporal events using DTM.

% figure 2
\begin{figure}
\centering
\includegraphics[height=2.5in, width=3.5in]{DTMfig1}
\caption{DTM topic 3 word prevalence over time. Date is YYYY-WW.}
\end{figure}

% figure 3
\begin{figure}
\centering
\includegraphics[height=2.5in, width=3.5in]{DTMfig2}
\caption{DTM topic 5 word prevalence over time. Date is YYYY-WW.}
\end{figure}

@@

 this will take time to complete (running download, prob 5 days, so approx April 19th)

1 figure (heatmap of users vs topics)

1 table (top words for 10 topics)

\subsection{Sentiment Classification}

% top positive and negative example tweets
% report accuracy
% (potentially) report topic modeling on top postive and negative PrEP tweets

Lastly we sought to create a classification scheme to classify the sentiment of tweets either positive or negative. This classifier would allow us to quickly identify positive and negative PrEP related tweets to guide public health efforts. We first performed Paragraph2Vec on both our PrEP related tweet corpus and on another tweet corpus that had binary sentiment labels, positive, or negative. Then we trained a simple logistic regression classifier on the paragraph-vectors. We found that our classifier had an accuracy of 70\% on a validation set.

We then used this classifier to classify our PrEP related tweets into positive or negative labels. We reported the most positive, and most negative tweets, by log(probability), on our full dataset, and on tweets that specifically mention either 'prep' or 'truvada' (Table 4). It seems that overall the classifier seems to capture the sentiment relatively well, however it clearly fails to recognise the sarcasm present in the most positive overall tweet.

The sentiment results aren't perfect, but they do help us identify positive individuals in PrEP and HIV public health such as user @greg0wen. The sentiment tweets also let us uncover some of the prevailing stigmas and negative sentiments surrounding PrEP usage.

% do a top positive tweets table

% do a top negative tweets table

@@

Acknowledgements (optional) go here.
